Evaluation of +S2 sea ice forecast and some other better title

### 2025-07-09 analysis/paper/access-ice.qmd
Simplifies hindcast-extent plot.

hindcast_bias_extent <- hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |> 
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |> 
  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset], 
            aes(x = as.Date(time),
                group = dataset)) + 
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
             aes(fill = month), 
             shape = 24, size = 2.2) +

  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
             shape = 21, size = 2) +

  scale_x_date(NULL, date_breaks = "1 month", 
               date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = labels_extent) +
  scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
                     aesthetics = c("fill", "color"), guide = "none") +
  facet_wrap(~ model, ncol = 1, labeller = labeller(model = labels_models)) +
  tag_facets()

hindcast_bias_extent_dif <- copy(hindcast_bias_extent)

hindcast_bias_extent_dif$layers[[1]] <- geom_hline(yintercept = 0) 
hindcast_bias_extent_dif$data <- extent_daily[hindcast_extent, on = c("time")] |> 
  _[, median_ci(i.aice - aice), 
    by = .(forecast_time = update(forecast_time, year = 2001),
           lag = as.numeric(as.Date(time) - forecast_time),
           model)] |> 
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)]

hindcast_bias_extent +
  hindcast_bias_extent_dif

>>>>

hindcast_extent[, .(model, forecast_time, member, time, Forecast = aice)] |> 
  extent_daily[i = _, on = c("time")] |> 
  _[, Difference := Forecast - aice] |> 
  _[, aice := NULL] |> 
  tidyfast::dt_pivot_longer(cols = c(Forecast, Difference)) |> 
  _[, median_ci(value), 
    by = .(dataset, 
           model, 
           measure = name,
           forecast_time = update(forecast_time, year = 2001),
           lag = as.numeric(as.Date(time) - forecast_time))] |> 
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |> 
  _[, model := factor(model, levels = c("S1", "S2"))] |> 
  _[, measure := factor(measure, levels = c("Forecast", "Difference"))] |>

  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset] |> 
              _[, measure := factor("Forecast", levels = c("Forecast", "Difference"))], 
            aes(x = as.Date(time),
                group = dataset)) +

  geom_hline(data = data.table(y = 0, measure = factor("Difference", levels = c("Forecast", "Difference"))), aes(yintercept = 0)) +
  geom_line(aes(colour = month, 
                group = interaction(forecast_time, time != time2))) +
  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], 
                           by = .(forecast_time, model, measure)], 
             aes(fill = lubridate::month(time2, label = TRUE)), 
             shape = 24, size = 2.2) +
  geom_point(data = \(x) x[lag == 1], 
             aes(fill = lubridate::month(time2, label = TRUE)), 
             shape = 21, size = 2) +

  scale_x_date(NULL, date_breaks = "1 month", 
               date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = labels_extent) +
  scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
                     aesthetics = c("fill", "color"), guide = "none") +
  facet_grid(measure ~ model, scales = "free_y")
```
